---
title: "Project proposal"
author: "Sarah Winne"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction

Has the average lexical richness (variety of vocabulary) changed over time? 

The data I'm using to answer this question is "Project Gutenberg creative literature life circumstance & work metadata collection"(Langer, Lars; 2023), found in the Harvard Dataverse. 
Each case is a unique book found in the Project Gutenberg digital library. There are 13493 cases in this dataset. The variables are the ID of the book, how many unique terms for organisms there are in each book (normalized to 10,000 tokens), the size of the vocabulary of each book, authors age at publication, the year of publication (or conception, if earlier), what type of literature the book is, gender of the author, whether or not the author is a parent, what the authors highest form of education was, whether the auhtor's occupation was connected to biodiversity, where the author lived most of their lives, how many reigons the author reached in their life, if the author's main residence was in a village, town, or city, then booleans for genre tags(satire, historical, social, children, adventure, mystery, crime, sci-fi, romance, fantasy, biography, travellogue). This is a total of 25 variables.

## 2. Data

```{r echo=FALSE}
data <- read_csv("ProjectGutenbergMetadata.csv", show_col_types = FALSE)

# I couldn't figure out how to access the csv file in the data folder from the proposal folder. I tried read_csv("data/ProjectGutenbergMetadata.csv"), but it said it didn't exist in the current working directory. So I just made another copy of the csv file within the proposal folder.
```

```{r echo=FALSE}
glimpse(data)
```

## 3. Data analysis plan

I will be using the year of publication to predict the lexical richness.

I will likely be comparing by year, or decade if that ends up working better. 



I expect lexical richness to have dropped in general, but there's a good chance it may have only dropped in certain genre groups or not at all. If it hasn't dropped, I will spend some more time comparing between genre groups or involving their highest level of education, as those seem most likely to affect the lexical richness. 