---
title: "Project proposal"
author: "Sarah Winne"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction

How has the average, maximum, and minimum lexical richness (variety of vocabulary) changed over time, especially around the 20th century, and how does the level of education or age of the author have an effect on the vocabulary used? 

The data I'm using to answer this question is "Project Gutenberg creative literature life circumstance & work metadata collection"(Langer, Lars; 2023), found in the Harvard Dataverse. 
Each case is a unique book found in the Project Gutenberg digital library. There are 13493 cases in this dataset. The variables are the ID of the book, how many unique terms for organisms there are in each book (normalized to 10,000 tokens), the size of the vocabulary of each book, authors age at publication, the year of publication (or conception, if earlier), what type of literature the book is, gender of the author, whether or not the author is a parent, what the authors highest form of education was, whether the auhtor's occupation was connected to biodiversity, where the author lived most of their lives, how many reigons the author reached in their life, if the author's main residence was in a village, town, or city, then booleans for genre tags(satire, historical, social, children, adventure, mystery, crime, sci-fi, romance, fantasy, biography, travellogue). This is a total of 25 variables.

## 2. Data

```{r echo=FALSE}
data <- read_csv("ProjectGutenbergMetadata.csv", show_col_types = FALSE)

# I couldn't figure out how to access the csv file in the data folder from the proposal folder. I tried read_csv("data/ProjectGutenbergMetadata.csv"), but it said it didn't exist in the current working directory. So I just made another copy of the csv file within the proposal folder.
```

```{r echo=FALSE}
glimpse(data)
```

## 3. Data analysis plan

I will be using the year of publication to predict the lexical richness.

I will likely be comparing by decade. 

```{r echo=FALSE}
data %>%
  mutate("century" = (floor(Year_Publication/100)*100)) %>%
  count(century)
```
Understanding how many publications that are in the Project Gutengerg collection from each century helps show how there are many more publications included that are in the 18-1900's, so my analysis might focus more where I have an abundance of data. 

```{r echo=FALSE}
data %>%
  mutate("decade" = (floor(Year_Publication/10)*10)) %>%
  ggplot(mapping = aes(x = Lexical_Richness, y = decade, alpha = .1)) +
  geom_point()
```

Graphing lexical richness against each decade gives a clearer look at how the general spread of the data is. 

```{r echo=FALSE}
data %>%
  mutate("decade" = (floor(Year_Publication/10)*10)) %>%
  ggplot(mapping = aes(x = Lexical_Richness, y = decade, alpha = .1)) +
  geom_point() +
  facet_wrap(~Highest_Education)
```

```{r echo=FALSE}
data %>%
  mutate("decade" = (floor(Year_Publication/10)*10)) %>%
  mutate("age_by_decade" = floor(Age_Publication/10)) %>%
  ggplot(mapping = aes(x = Lexical_Richness, y = decade, alpha = .1)) +
  geom_point() +
  facet_wrap(~age_by_decade)
```

These two faceted graphs, by the highest education reached and the author's age(rounded by decade) respectively, helps highlight a few outliers within each fascet. For example the point in the academic facet on the 1900 line, to the far right, showing by far the highest lexical richness.
The second graph shows which ages had the most authors and the widest spread of lexical richness. It would be interesting to see the avgerages, but also the maximums and minimums of each age group, and see if there had been similar or different changes in each age group over time. 

It might be useful to use proportions or densities more, so it's easier to see the differences or similarities between facets with a lot of data, or less data (both in the un-faceted and faceted graphs).
Some more summary statistics about the average, max, and min of each decade for lexical richness would be useful. Also for the faceted groups.

To support my hypothesis that average lexical richness was decreasing, I would need to see the average go down over the increasing decades. 
