---
title: "Lexical Richness Over The 17th-20th Centuries"
subtitle: "Data from Project Gutenberg"
author: "Sarah Winne"
institute: "Houghton University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
gutenberg <- read_csv("/cloud/project/data/ProjectGutenbergMetadata.csv")
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "/Users/sarahwinne/Documents/Houghton/DATA/BookBG.jpg"
)
```

class: center, middle

## How has lexical richness changed over the 17th-20th centuries?

---

# Lexical Richness

- The variety of vocabulary used

- High lexical richness does not automatically make a book good

--

- Well known books like "To Kill a Mockingbird" or "Of Mice and Men" have simpler vocabularies

- "Moby-Dick" and "The Book Thief" have much more extensive vocabularies

---

class: inverse, center, middle

# Limitations

---

# Gutenberg Project

.pull-left[
- Online archive of free e-books

- Mostly older literary works with expired copyrights (in the US)

- Lack of recent works 

- Less 17th-18th century works than 19th-20th century works

]
.pull-right[
```{r recode_gutenberg, echo = FALSE}
gutenberg <- gutenberg %>%
  mutate("century" = (floor(Year_Publication/100)*100)) %>%
  mutate("decade" = (floor(Year_Publication/10)*10)) %>%
  mutate("age_by10" = (floor(Age_Publication/10)*10))
```

```{r plot_gutenberg1, warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults to make the figure bigger
gutenberg %>%
  mutate(Year_Publication = if_else(Year_Publication<=1800, Year_Publication, NA)) %>%
  ggplot(mapping = aes(x = as_factor(century))) +
  geom_bar()+
  theme_minimal() +
  labs(title = "Gutenberg's Collection",
       x = "Century Published",
       y = "E-Book Count")
```
]

---

class: inverse, center, middle

# Lexical Richness Changes

---

```{r plot-gutenberg2, echo = FALSE, message = FALSE}
gutenberg %>%
  ggplot(aes(y = Lexical_Richness, x = Year_Publication, color = century, alpha = .1)) +
  geom_point() +
  geom_smooth(color = "black", method = "lm", se = FALSE) + 
  guides(alpha = "none", color = "none") +
  theme_minimal() + 
  labs(title = "Overall Decrease in Lexical Richness over the Years", 
       subtitle = "1635-1969",
       x = "Year Published", 
       y = "Lexical Richness")
```

---

```{r plot-gutenberg3, warning=FALSE, echo = FALSE, message = FALSE}
gutenberg %>%
  ggplot(aes(y = Lexical_Richness, x = Year_Publication, color = century, alpha = .1)) +
  geom_point() +
  geom_smooth(color = "black", se = FALSE) + 
  guides(alpha = "none", color = "none") +
  theme_minimal() + 
  labs(title = "Lexical Richness over the Years", 
       subtitle = "1635-1969",
       x = "Year Published", 
       y = "Lexical Richness")
```

---

# Range

.pull-left[

- Median lexical richness decreases

- Upper range of lexical richness increases

- Overall quantity of e-books increases
]

.pull-right[
```{r plot-gutenberg4, warning=FALSE, out.width="100%", fig.width=4, echo=FALSE, message = FALSE}
gutenberg %>%
  ggplot(mapping = aes(y = Lexical_Richness, x = decade, group = century, alpha = 0.1)) +
  geom_boxplot() +
  theme_minimal() + 
  guides(alpha = "none") +
  labs(title = "Lexical Richness",
       subtitle = " Split by Centuries", 
       x = "Year Published", 
       y = "Lexical Richness")
```
]

---

```{r plot-gutenberg5, warning=FALSE, echo = FALSE, message = FALSE}
gutenberg %>%
  ggplot(mapping = aes(y = century, fill = Literature_Form, color = Literature_Form)) +
  geom_bar(position = "fill") +
  labs(title = "Literature Forms Proportions",
       subtitle = " Split by Centuries", 
       x = "Proportion", 
       y = "Century")
```

---

```{r plot-gutenberg6, warning=FALSE, echo = FALSE, message = FALSE}
gutenberg %>%
  ggplot(aes(y = Lexical_Richness, x = Year_Publication, color = century, alpha = .1)) +
  geom_point() +
  geom_smooth(color = "black", se = FALSE) + 
  guides(alpha = "none", color = "none") +
  theme_minimal() + 
  labs(title = "Lexical Richness over the Years", 
       subtitle = "1635-1969",
       x = "Year Published", 
       y = "Lexical Richness")
```

---
